A minimal reaction-diffusion neural model generates {\emph{C. elegans}} undulation

### 2020-06-07 paper.tex
work
more work

% TODO confirm the Xu reprint policy
% TODO show how the latest paper describes a small sequence of neurons forming coupled oscillators.

@@

In more complex animals, like horses or dogs, there is only one CPG; the dynamics of the CPG change depending on the gait of the animal \citep{collins1994}.  The mode of locomotion depends on the sequencec in which the neurons fire, but the network's topology remains the same.  We show later that breaking the symmetry of the CPG model can lead to radical changes in behaviour.

\begin{figure}[h!]
    \label{fig: xu_cpg}
    \centering
    \includegraphics[width=9cm]{figures/xu_cpg/xu_cpg.png}
    \caption{Pirated from Xu.}
\end{figure}

@@

\begin{equation} % reference HH model here as well
    \label{eq: fhn}
    \begin{aligned}
        dv &= f(v) + w - I_\mathrm{ext}\\
        dw &= \epsilon (a - bv)\\
        f(v) &= \frac{v^3}{3} - v
    \end{aligned}
\end{equation}

@@

% TODO: remove Morris-Lecar
% remark that if the w and v nullclines come near each other, you see a critical slowdown.
% In Hodgkin-Huxley notation, $n$ is the potassium channel.  In the FHN formulation, this corresponds to the $w$ gate variable.
\begin{figure}[h!]
    \label{fig: nm nullclines}
    \centering
    \inputtikz{figures/neuron_dynamics/neuron_dynamics}
    % TODO: change legend to dv/dt=0, dw/dt=0
    % relabel ``real c. elegans'' to C. elegans (experimental data).
    % for publication: remove ticklabels, keep minor grid
    % Replace the figure with Xu's
    % note that in the vicinity of the v-nullcline, the w-nullcline is essentially straight.
    \caption{Nullclines of several different neuronal models; on the right is the biological model of Xu et al, in the centre is the Morris-Lecar model, and to the right is the FitzHugh-Nagumo model.  These have been arranged in order of decreasing complexity.}
\end{figure}

@@

\begin{figure}[h!]
    \label{fig: fhn_dynamics}
    \centering
    \includegraphics{figures/fhn_dynamics/fhn_dynamics.pdf}
    % TODO describe this more.
    % Show a sample trajectory in each of these.
    % remove the ticklabels
    % condense into single figure
    % .
    \caption{The nullclines of the FitzHugh-Nagumo neuron.  Nullclines are isoclines where the derivative of a variable is zero.  Here we show an oscillatory mode and an excitable one.  In an oscillating cell, the stable equilibrium is replaced by an unstable equilibrium (open circle) around which it circles indefinitely.  See \citet{parsons2018} for a clear explanation of the dynamics of oscillatory biological systems.}
\end{figure} % TODO: legend

@@

\begin{figure}[h!]
    \centering
    \pgfplotsset{}
    \inputtikz{figures/anal_sim_comp/anal_sim_comp}
    \caption{Comparison between simulated dynamics (top row) and observed dynamics in circuit (bottom row).}
    \label{fig: anal_sim_comp}
\end{figure}

@@

Our model of the nematode is surprisingly similar to real unconstrained motion.

% Can we use this to explore how mutations effect motion, and mechanisms behind differences in phenotypes.
% The particular model doesn't really matter; if you have the right topology, you can generate the behaviour.
% Reservoir computing and our neurons.  Speculate that different circuits of FHN neurons might be suitable dynamical reservoirs.  However, a CPG has a range of comparatively simple dynamics.
% cite Sanchez-Sinencio, Spano on circuits.  Theirs emphasize accuracy whereas ours emphasizes speed and minimal cost.

>>>>

% TODO show how the latest paper describes a small sequence of neurons forming coupled oscillators.

@@

In more complex animals, like horses or dogs, there is only one CPG; the dynamics of the CPG change depending on the gait of the animal \citep{collins1994}.  The mode of locomotion depends on the sequence in which the neurons fire, but the network's topology remains the same.  We show later that breaking the symmetry of the CPG model can lead to radical changes in behaviour.

\begin{figure}[h!]
    \label{fig: xu_cpg}
    \centering
    \includegraphics[width=9cm]{figures/xu_cpg/xu_cpg.png}
    \caption{The central pattern generator proposed by \citet{xu2018}.  Note the structure of the neurons, especially the head oscillator driving the descending pathway.}
\end{figure}

@@

\begin{equation}
    \label{eq: fhn}
    \begin{aligned}
        dv &= f(v) + w - I_\mathrm{ext}\\
        dw &= \epsilon (a - bv)\\
        f(v) &= \frac{v^3}{3} - v
    \end{aligned}
\end{equation}

The parameter $I_\mathrm{ext}$ is an external driving current, and can also be used to increase excitability.  In the analog implementation, this is an integral component of tuning the neuron's frequency response.

@@

% TODO: remove Morris-Lecar
% remark that if the w and v nullclines come near each other, you see a critical slowdown.
% In Hodgkin-Huxley notation, $n$ is the potassium channel.  In the FHN formulation, this corresponds to the $w$ gate variable.
\begin{figure}[h!]
    \label{fig: nm nullclines}
    \centering
    \includegraphics{figures/neuron_dynamics/neuron_dynamics.pdf}
    % note that in the vicinity of the v-nullcline, the w-nullcline is essentially straight.
    \caption{Nullclines of several different neuronal models; on the right is the biological model of Xu et al, in the centre is the Morris-Lecar model, and to the right is the FitzHugh-Nagumo model.  These have been arranged in order of decreasing complexity.}
\end{figure}

@@

\begin{figure}[h!]
    \label{fig: fhn_dynamics}
    \centering
    \includegraphics{figures/fhn_dynamics/fhn_dynamics.pdf}
    \caption{The nullclines of the FitzHugh-Nagumo neuron.  Nullclines are isoclines where the derivative of a variable is zero.  Here we show an oscillatory mode and an excitable one.  In an oscillating cell, the stable equilibrium is replaced by an unstable equilibrium (open circle) around which it circles indefinitely.  See \citet{parsons2018} for a clear explanation of the dynamics of oscillatory biological systems.}
\end{figure} % TODO: legend

@@

\begin{figure}[h!]
    \centering
    \pgfplotsset{}
    % \inputtikz{figures/anal_sim_comp/anal_sim_comp}
    \includegraphics{figures/anal_sim_comp/anal_sim_comp}
    \caption{Comparison between simulated dynamics (top row) and observed dynamics in circuit (bottom row).}
    \label{fig: anal_sim_comp}
\end{figure}

@@

Our model of the nematode is surprisingly similar to real unconstrained motion.  While most measurements of \emph{C. elegans} are taken on horizontal agar sheets, observations made in clear, water-filled cuvettes where the worm floats freely seem to resemble our simulated worm more.

Given the flexibility of the model, by introducing defects and asymmetries into the topology, we can study how mutations affect the nematode's motion, as well as the structures underlying differences in phenotypes.

% The particular model doesn't really matter; if you have the right topology, you can generate the behaviour.
% Reservoir computing and our neurons.  Speculate that different circuits of FHN neurons might be suitable dynamical reservoirs.  However, a CPG has a range of comparatively simple dynamics.
% cite Sanchez-Sinencio, Spano on circuits.  Theirs emphasize accuracy whereas ours emphasizes speed and minimal cost.
